\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-10-01 NAR_DRAFT.tex
I put the frequency in there. check results

\rngcomment{Need paragraph in discussion emphasizing importance of calibration.}
The improved calibration of \batcave provides several advantages, including a tumor-specific means to select a threshold for positive classification that exhibits interpretable positive predictive value properties.
Cancer genomics for research and drug discovery is an important yet costly endeavour, and manual review of variants is often required to limit the number of false leads followed.
Improved calibration allows researchers to select a tolerable false positive rate, and focus efforts on a set of variants with known probability of being present.
In the clinic, identified variants act as biomarkers for susceptibility to targeted drugs, and well calibrated probabilities facilitate the use of risk models in the choice of treatment rather than an all or nothing approach.
\bkmcomment{I mean by "risk models", if you can assign a probability of the mutation being present you can make informed tradeoffs between likelihood of benefit and risk of harm for a particular targeted therapy.}
A properly calibrated model implies a known rate of false positives.
One reason why tumors are not routinely sequenced to high depth is that as sequencing depth increases, false positive rate increases without bound \citep{Cibulskis2013}.
The International Cancer Genome Consortium recommends that catalogues of somatic mutation target a true positive rate of 95\%, and an acquisition rate of 80\% of all mutations present in a sample \texttt{https://icgc.org/icgc/goals-structure-policies-guidelines/e8-genome-analyses}.
Assuming a sample size of 1 million cells, and a per-base mutation rate of $3 \dot 10^{-8}$, implies 10 million mutations per sample in a neutrally evolving tumor.
A naive application of Eqn. \ref{eqn:mut_rate} implies a minimum variant allele frequency of \bkmcomment{calculate this, will be very low.}, requiring sequencing depths far in excess of current standards.
Proper calibration will allow deeper sequencing without ballooning false positive rates.
\bkmcomment{we can talk about what should actually survive from this previous section. I like the specific example but may be too wordy here. One of the things I have tried to do is establish that ppv qualities arise from superior calibration, and that allows the use of calibration as short hand here in the discussion. This may not be well done.}

>>>>

\rngcomment{Need paragraph in discussion emphasizing importance of calibration.}
The improved calibration of \batcave provides several advantages, including a tumor-specific means to select a threshold for positive classification that exhibits interpretable positive predictive value properties.
Cancer genomics for research and drug discovery is an important yet costly endeavour, and manual review of variants is often required to limit the number of false leads followed.
Improved calibration allows researchers to select a tolerable false positive rate, and focus efforts on a set of variants with known probability of being present.
In the clinic, identified variants act as biomarkers for susceptibility to targeted drugs, and well calibrated probabilities facilitate the use of risk models in the choice of treatment rather than an all or nothing approach.
\bkmcomment{I mean by "risk models", if you can assign a probability of the mutation being present you can make informed tradeoffs between likelihood of benefit and risk of harm for a particular targeted therapy.}
A properly calibrated model implies a known rate of false positives.
One reason why tumors are not routinely sequenced to high depth is that as sequencing depth increases, false positive rate increases without bound \citep{Cibulskis2013}.
The International Cancer Genome Consortium recommends that catalogues of somatic mutation target a true positive rate of 95\%, and an acquisition rate of 80\% of all mutations present in a sample \texttt{https://icgc.org/icgc/goals-structure-policies-guidelines/e8-genome-analyses}.
Assuming a sample size of 1 million cells, and a per-base mutation rate of $3 \dot 10^{-8}$, implies 10 million mutations per sample in a neutrally evolving tumor.
A naive application of Eqn. \ref{eqn:mut_rate} implies a minimum variant allele frequency of around .00001, requiring sequencing depths far in excess of current practice.
Proper calibration will allow deeper sequencing without ballooning false positive rates.
\bkmcomment{we can talk about what should actually survive from this previous section. I like the specific example but may be too wordy here. One of the things I have tried to do is establish that ppv qualities arise from superior calibration, and that allows the use of calibration as short hand here in the discussion. This may not be well done.}
